---
title: "TCGA_analysis_GF_082824"
output: html_document
date: "2024-08-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script uses processed data from TCGA_analysis_data_clean_prep_Fort.R
## script to perform survival and correlation analyses

## These analyses are related to Figures 1A, S1A, S2A, and S2J


```{r}
# Load required packages
library(ggplot2)
library(data.table)
library(dplyr)
library(survival)
#install.packages('survminer')
library(survminer)
#install.packages('tab')
library(tab)
#install.packages('arsenal')
library(arsenal)
library(ggcorrplot)

# Set working directory 
setwd('/uufs/chpc.utah.edu/common/home/snydere-group2/Gabby/KPH_TCGA_analysis')

# Read in data from data preparation/cleaning script
tcga_data = read.csv("/uufs/chpc.utah.edu/common/home/snydere-group2/Gabby/KPH_TCGA_analysis/TCGA_KPH_survival_HNFstatus.csv")

```


## HNF4A survival analysis - related to Figure 1A

### KM Curve
```{r, echo=TRUE}
survival_fit = survfit(Surv(Overall_Survival_Months,OS_status) ~ HNF_status, data = tcga_data)
ggsurvplot(survival_fit, conf.int=F, pval = T, pval.method = F, risk.table = F, legend.title = "HNF4a Status",
           legend.labs = c("HNF4a-High","HNF4a-Low"),palette = c('maroon','darkcyan'))
```

### Run coxPH model 
```{r}
cox = coxph(Surv(Overall_Survival_Months,OS_status) ~ HNF_status, data=tcga_data) 
summary(cox)
```

### Make table to visualize distribution of various potential confounding variables
```{r, results="asis"}
concordance_table <- tableby(HNF_status ~ Diagnosis_Age + Ethnicity_Category + Race_Category + Sex + Tumor_Type + American_Joint_Committee_on_Cancer_Metastasis_Stage_Code +
                              American_Joint_Committee_on_Cancer_Tumor_Stage_Code +
                              Neoplasm_Disease_Stage_American_Joint_Committee_on_Cancer_Code, data=tcga_data)
summary(concordance_table)
```


## EPS8L3 survival analysis - related to Figure S1A

### Looking at top 47 expressors vs all others (same n's as HNF4A-high vs HNF4A-low groups)
```{r}
tcga_data = tcga_data %>% dplyr::mutate(EPS8L3_top = dplyr::case_when(EPS8L3 > quantile(EPS8L3,0.88) ~ "High",
                                                                               TRUE ~ "Low"))

survival_fit = survfit(Surv(Overall_Survival_Months,OS_status) ~ EPS8L3_top, data = tcga_data)
ggsurvplot(survival_fit, conf.int=F, pval = T, pval.method = F, risk.table = F, legend.title = "EPS8L3 Level",
           legend.labs = c("High","Low"),palette = c('darkcyan','maroon'))
```

### Run coxPH model 
```{r}
cox = coxph(Surv(Overall_Survival_Months,OS_status) ~ EPS8L3_top, data=tcga_data) 
summary(cox)
``` 


## Correlation Analysis of HNF4A with other genes of interest - related to Figure S2A

### Scatterplots
```{r}
genes_oi = c('LGALS4','HNF1A','EPS8L3')

for (gene in genes_oi) {
  print(ggscatter(tcga_data, x='HNF4A',y=gene,color='maroon',cor.coef = T,cor.method='spearman',
                  conf.int=TRUE,add = c("reg.line"),ylab = paste0(gene," log2(TPM+1)"), xlab = "HNF4A log2(TPM+1)",
                  title = paste0("Correlation between HNF4A and ",gene)))
}
```


## Correlation Analysis of HNF4A with ssGSEA scores - related to Figure S2J

### Scatterplots ('3154_1651_combined' score shown in manuscript)
```{r}
scores_oi = colnames(tcga_data[74:80])

for (score in scores_oi) {
  print(ggscatter(tcga_data, x='HNF4A',y=score,color='maroon',cor.coef = T,cor.method='spearman',
                  conf.int=TRUE,add = c("reg.line"),ylab = score, xlab = "HNF4A log2(TPM+1)",
                  title = "Correlation between HNF4A and ssGSEA scores"))
}
```







